---
title: "Project for Getting and Cleaning data course"
author: "GM"
date: "26/07/2015"
output:
  html_document:
    keep_md: yes
---

## Project Description
The purpose of the project is to demonstrate ability to collect, work with, and clean a data set.

##Study design and data processing

###Collection of the raw data
The data was obtained from the following url:
https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip 
It was then unzipped and saved in a folder.

###Notes on the original (raw) data 
The data represent data collected from the accelerometers from the Samsung Galaxy S smarthpone.

##Creating the tidy datafile

###Guide to create the tidy data file
The script "run_analysis.R" carries out the full process from downloading the data to producing the final tidy dataset.
To download and run the script run the following code in R, ensuring that the script is in your current working directory: 
```{r}
        source("run_analysis.r")
        run_analysis.r()
```

###Cleaning of the data
The cleaning script is called "run_analysis.R" and is saved in the main data folder described above.

The script does the following:
* There are 8 different files that need to be read and transformed into data tables
* From these, 6 are merged into a single data set. Namely, 
        + the 2 files corresponding to the training and test measurements; 
        + the 2 files corresponding to the training and test activities, and
        + the 2 files corresponding to the subjects who performed each activity for the training and test sets.
* The "features.txt" file is used to name the columns of the merged dataset.
* Once merged, only the columns which correspond to the measurements on the mean and standard deviation are extracted.
* The activities performed by each subject are represented by numbers in the data set. The file "activity_labels.txt" is used to substitute the numbers for their corresponding descriptive labels (e.g. 1 substituted for WALKING)
* Descriptive names are given to each of the labels in the data set
* A new data set is created which contains the average of each variable in the clean dataset for each activity and each subject.

A detailed description of the script run_analysis.R can be found here:
https://github.com/shinosuke85/Getting-and-Cleaning-Data/blob/master/README.md

##Description of the variables in the tidy_data.txt file
General information:
 * Dimensions of the dataset: 180 rows x 68 columns
 * Summary of the data: use the code in R
 to obtain the range and quartiles of each of the variables 
  ```{r}
  table <- read.table("tidy_data.txt")
  summary(table)
  ```
 * Variables present in the dataset: 68 variables (described below)
 
### Variable 1: subjectid
Each row identifies the subject who performed the activity for each window sample.
 + Class: integer
 + Unique values/levels of the variable: values 1 to 30.
 + Unit of measurement: none

### Variable 2: activity
Each row identifies which activity was performed when the measurements were taken
 + Class: factor
 + Unique values/levels of the variable: 6 levels - WALKING, WALKING_UPSTAIRS, WALKING_DOWNSTAIRS, SITTING, STANDING, LAYING
 + Unit of measurement: none

## Variables 3 to 68: measurements
Each row shows the measurement values obtained using a smartphone. 
 + Class: double
 + Unique values/levels of the variable: all measurements are normalised and bounded within [-1,1].
 + Unit of measurement: none

Each measurement is described by the following:
1. Type of signal:
        + t := time domain signal
        + f := frequency domain signal
2. Acceleration type:
        + BodyAcc := Body acceleration signal
        + GravityAcc := Gravity acceleration signal
        + BodyGyro := Angular velocity
3. Jerk := Jerk signal derived from measures above
4. Mag := used when measurement indicates the magnitude of the signal, calculated using the Euclidean norm.
5. 3-axial signals: estimates of the feature vector for each pattern are calculated. X, Y and Z are used to indicate the direction of the measurement
6. Estimates:
        + mean:= mean value
        + standarddeviation := Standard deviation


##Sources
Data description: http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones
